---
title: "Individual Assignment 3"
author: "Ali Rafieepouralavialavijeh"
date: "3/13/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(knitr)    # For knitting document and include_graphics function
library(gridExtra) # For including grids of plots
library(psych)
library(ggiraph)
library(ggiraphExtra)
library(car)
library(GGally) #for ggpairs
library(lubridate)

```

Loading the data

```{r}
data <- read.csv("C:\\Users\\urafi\\OneDrive\\Desktop\\csv3.csv")
data_annual <- read.csv("C:\\Users\\urafi\\OneDrive\\Desktop\\csv2.csv")
```

checking the data within the dataset!

```{r}
head(data)
head(data_annual)
```

edit year


```{r}
ggplot(data, aes(y=complaints, x=month.year)) +
  geom_point() +
  labs(x="Consectutive months from 2011 untill 2021",
         y="Number of complaints recived per month");

```


```{r}
ggplot(data_annual, aes(y=X0, x=Week.Year)) +
  geom_point() +
  labs(x="Consectutive months from 2011 untill 2021",
         y="Number of complaints recived per month");

```



```{r}
summary(lm(complaints ~ month.year, data=data))
summary(lm(X0 ~ Week.Year, data=data_annual))

```

Dropping the outlier

```{r}
reg <- slice(data, -c(112))
reg.model = lm(complaints ~ month.year, data=reg)
summary(reg.model)

ggplot(reg, aes(y=complaints, x=month.year)) +
  geom_point() +
  labs(x="Consectutive months from 2011 untill 2021",
         y="Number of complaints recived per month");

```

All three predictor variables have an influence on influence on albums sales at the 5% level of significance: adverts (t(196)=12.261, p < 2e-16), airplay (t(196)=12.123, p < 2e-16), and attract (t(196)=4.548, p < 9.49e-06). The intercept is not significantly different from 0 (t(196) = -1.534, p=0.127). R^2 is 0.6647, and adjusted R^2 is 0.6595. Coefficients and 95% confidence intervals are listed below.


```{r}
confint(reg.model)
```


```{r}
ggPredict(reg.model, se=TRUE)
reg
```

```{r}
durbinWatsonTest(reg.model)
plot(reg.model)
```



```{r}
reg$fitted<- reg.model$fitted
reg$res <- reg.model$residuals
reg$standardized.residuals <- rstandard(reg.model)


possible.outliers <- subset(reg, standardized.residuals < -1.96 | standardized.residuals > 1.96)

possible.outliers

```
We found 5 residuals are above or below 1.96 standard deviations. As this represents 4% of the observations, expected if the residuals are normal (5% of data is expected to be outside of 2 standard deviations), we do not consider any of these observations as outliers and continued with all 115 observations included in the model.



```{r}
reg$cooks <- cooks.distance(reg.model)
plot(sort(reg$cooks, decreasing=TRUE))
```

To investigate influential cases, we calculated Cook’s Distance on the developed model. Cook’s distance was a maximum of 0.0707659, far below the chosen cutoff value of 1. We thus conclude that there are no influential cases


```{r}
shapiro.test(reg.model$residuals)
```








VIF multico

```{r}
vif(reg.model)
```

predicting

```{r}
newValues=tibble(month.year=c(103, 104, 105,106,107,108,109,110,111,112,113,114,115,116,117,118,119,120,121,122,123,124,125,126,127,128,129))
predict.lm(reg.model, newValues)
```



deleting certain points

```{r}

data <- read.csv("C:\\Users\\urafi\\OneDrive\\Desktop\\csv3.csv")
ggplot(data, aes(y=complaints, x=month.year)) +
  geom_point() +
  labs(x="Consectutive months from 2011 untill 2021",
         y="Number of complaints recived per month");

sth = lm(complaints ~ month.year, data=data)
ggPredict(sth, se=TRUE)


```
